---
title: "Tutorial of mesounitR"
author: "David Farò"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial of mesounitR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mesounitR)
library(sf)
library(ggplot2)
library(dplyr)
```

Load sample data
```{r}
convert_points_to_voronoi <- function(flow_pts,depth_min) {

  voronoi_polygons <- sf::st_voronoi(st_union(flow_pts))
  
  flow_voronoi <- sf::st_collection_extract(voronoi_polygons, type = "POLYGON") %>%
    st_sf()
  
  # join original points (flow_pts) and voronoi polygons (flow_voronoi) to add attributes from original file
  flow_joined <- st_join(flow_voronoi,flow_pts)
  
  # select only wetted points, based on depth_min value
  flow_voronoi_wet <- flow_joined %>%
    filter(DEPTH >= depth_min)
  
  # convert sf to df
  flow_df <- as.data.frame(st_coordinates(flow_pts))
  
  ## compute convex hull. it will be filter out points outside of original hull
  # Get convex hull indices
  hull_idx <- chull(flow_df$X, flow_df$Y)
  
  # Close the polygon by repeating the first point
  hull_coords <- rbind(flow_df[hull_idx, c('X','Y')], flow_df[hull_idx[1], c('X','Y')])
  
  # Create polygon
  hull_polygon <- st_polygon(list(as.matrix(hull_coords)))
  hull_sf <- st_sfc(hull_polygon, crs = st_crs(flow_pts)) |> st_sf()
  
  ## select polygons within hull
  flow_wet <- flow_voronoi_wet[st_within(flow_voronoi_wet, hull_sf, sparse = FALSE)[,1], ]
  
  return(flow_wet)
}

# threshold to differentiate wet and dry elements
depth_min <- 0.05

# load points
flow_pts <- sf::st_read('data/sample_Mareta.shp')

# convert points to Voronoi polygons
flow_wet <- convert_points_to_voronoi(flow_pts,depth_min)

# clean out non-contiguous elements
flow_wet <- mesounitR::clean_mesh(flow_wet)
```

Step 0) Retrieve empirical parameter for Global Score 
```{r}
#### compute empirical parameters from surveys

## compute wetted channel widths --> used later during step 3
# load centerline
centerline <- st_read('data/centerline.shp')

## dissolve the flow mesh into 1 polygon for the wetted channel
# paremeter for holes to be filled (max_hole_area) might need to be adjusted
channel_wet <- mesounitR::flow_to_wetchannel(flow_wet,depth_min,20)

## compute average wetted channel width
# --> parameters longitudinal distance (trans_longidst) and transect width (trans_width) might have to adjusted
# --> calling function also plots transects, to give an idea if computation parameters should be adjusted
W <- mesounitR::compute_widths(channel_wet,centerline,20,300)

## load HMU maps
hmu_map <- st_read('data/sample_Mareta_survey.shp')

# if more maps availabe, computation should be repeated for each map (since W might change with changing discharge)
c_params <- mesounitR::compute_empirical_mesosize(hmu_map,W)
```

Step 1) (over-)segment mesh into superpixels (super-cells ?)
```{r}
# define average supercell area
area_supercell <- 50 # in m^2, or same unit as mesh
# define compactness parameter of "SLIC" algorithm
compactness <- 1000

# compute supercells
supercells <- mesounitR::elements_to_supercell(flow_wet,area_supercell,compactness)

supercells %>%
  ggplot() +
  geom_sf(aes(fill=factor(cluster)),color=NA) +
  theme_bw() +
  theme(legend.position = 'none')

#### write SP map
# st_write(supercells,'supercells_area50.shp',delete_dsn=T)
```
step 2) regionalize superpixels to a mosaic composed of n units using SKATER spatial clustering algorithm (from spdep package)
```{r}
#### define computation range
# number of potential regions in computed mosaics ranges from n_min to n_max, with increments defined by n_by
n_min <- 10
n_max <- 100
n_by <- 10

range_n <- seq(from=n_min,to=n_max,by=n_by)

#### compute potential mosaics using SKATER algorithm
tic()
list_units <- mesounitR::supercell_to_units(supercells,range_n)
toc()
```

step 3) Optimal segmentation
Compute optimality metrics (v,MI,ms) and the derived Global Score (GS)
--> select optimal region mosaic based on computed GS
```{r}
# compute optimality metrics
optimality_metrics <- compute_optimality_metrics(flow_wet,list_units,range_n,c_params)

p <- plot_segmentation_indicators(optimality_metrics)
p

# select optimal number of regions (and regionalization)
opt_n <- optimal_n(optimality_metrics)
opt_units <- list_units[[opt_n$i]]

st_write(opt_units,paste0('optimal_mosaic_n',opt_n$n,'.shp'),delete_dsn=T)
```


step 2 bis) Refine GS curve, by adding additional mosaics with n units to the list
```{r}
# ## 
range_n_toadd <- seq(25,65,by=10)
list_units_old <- list_units
add_units_results <- add_units(list_units_old,supercells,range_n_toadd)

range_n <- add_units_results$range_n
list_units <- add_units_results$list_units
```

