---
title: "Tutorial of mesounitR"
author: "David Farò"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial of mesounitR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sf)
library(ggplot2)
library(dplyr)
library(mesounitR)
```

Install mesounitR
```{r}
# install package from github
# install.packages("pak")
# pak::pak("david-faro/mesounitR")

# OR
# remotes::install_github("david-faro/mesounitR")

# load library
# library(mesounitR)
```

Load sample data
```{r}
# threshold to differentiate wet and dry elements
depth_min <- 0.05

# load hydraulic simulation
flow <- sf::st_read('data_tutorial1/flow_DV.shp')

# extract only wet polygons
flow_wet <- flow %>%
  filter(DEPTH >= depth_min)

# clean out non-contiguous elements
flow_wet <- clean_mesh(flow_wet)
```

Step 0) Retrieve empirical parameter for Global Score
```{r}
# load centerline
centerline <- st_read('data_tutorial1/centerline.shp')

## load surveyed mesohabitat map
hmu_map <- st_read('data_tutorial1/survey_hmu.shp')

## dissolve flow_wet or surveyed HMU polygons into 1 polygon for the wetted channel
# adjust max_hole_area if needed
channel_wet <- flow_to_wetchannel(hmu_map,max_hole_area = 20) # using flow_wet, since modelled area larger than surveyed one

# alternatively: use flow_wet if reach extent and discharge same as that of surveyed hmu
# channel_wet <- flow_to_wetchannel(flow_wet,max_hole_area = 20)

## compute average wetted channel width
# adjust parameters trans_longidst and trans_width if needed
W_emp <- compute_widths(channel_wet,centerline,20,300)

# compute parameters
c_params <- compute_empirical_mesosize(hmu_map,W_emp)
```

Step 1) (over-)segment mesh into supercells
```{r}
# define average supercell area
area_supercell <- 50 # in m^2, or same unit as mesh
# define compactness parameter of "SLIC" algorithm
compactness <- 300

# compute supercells
supercells <- elements_to_supercell(flow_wet,area_supercell,compactness)

# plot supercells
supercells %>%
  ggplot() +
  geom_sf(aes(fill=factor(cluster)),color=NA) +
  theme_bw() +
  theme(legend.position = 'none')

#### write SP map
# st_write(supercells,'supercells.shp',delete_dsn=T)
```
step 2) regionalize superpixels to a mosaic composed of n units using SKATER spatial clustering algorithm (from spdep package)
```{r}
#### define computation range
# number of potential regions in computed mosaics ranges from n_min to n_max, with increments defined by n_by
n_min <- 5
n_max <- 65
n_by <- 10

n_range <- seq(from=n_min,to=n_max,by=n_by)

#### compute potential mosaics using SKATER algorithm
units_res <- supercell_to_units(supercells,n_range)

list_units <- units_res$list_units # list of mosaics
skater_res <- units_res$skater_res # skater object, contains information about all possible segmentations with number of units n = 1 ... N
```

step 3) Optimal segmentation
Compute optimality metrics (v,MI,ms) and the derived Global Score (GS)
--> select optimal region mosaic based on computed GS
```{r}
# compute wetted channel area of modelled reach
A_tot <- as.numeric(sum(st_area(flow_wet)))

# update channel width and W_area based on modelled reach
channel_wet <- flow_to_wetchannel(flow_wet,max_hole_area = 20)

## compute average wetted channel width
# adjust parameters trans_longdist and trans_width if needed
W_mod <- compute_widths(channel_wet,centerline,20,300)

c_params$W_area <- W_mod^2

# compute optimality metrics
optimality_metrics <- compute_optimality_metrics(flow_wet,list_units,n_range,c_params,A_tot)

plot_segmentation_indicators(optimality_metrics)

# select optimal number of regions (and regionalization)
opt_n <- optimal_n(optimality_metrics)
opt_units <- list_units[[opt_n$i]]

print(paste0('optimal n is ',opt_n$n))

# write optimal shp
# st_write(opt_units,paste0('optimal_mosaic_n',opt_n$n,'.shp'),delete_dsn=T)
```


step 2 bis) Refine GS curve, by adding additional mosaics with n units to the list
```{r}
# add mosaics with n units defined in n_range_toadd 
n_range_toadd <- c(23,24,26,27,28,29)

list_units_old <- list_units
add_units_results <- add_units(list_units_old,skater_res,supercells,n_range_toadd)

n_range <- add_units_results$n_range
list_units <- add_units_results$list_units

```
--> re-run step 3 for updated results
